---
title: "p8105_hw5_zx2482"
author: "Allison Xia"
date: "2023-11-15"
output: github_document
---
```{r}
library(dplyr)
library(tidyverse)
library(stringr)
library(broom)
library(purrr)
library(ggplot2)
```

### Problem 1

```{r load data}
homicides = read.csv("homicide-data.csv")
str(homicides)
```

#### Describe the raw data.

```{r}
homicides =  
  homicides |> 
  unite("city_state", city:state, sep = ", ", remove = FALSE)
homicides |> 
  group_by(city) |> 
  summarize(total_count = n())
homicides |> 
  filter(disposition == "Closed without arrest" | disposition =="Open/No arrest") |> 
  group_by(city) |> 
  summarize(count = n())
```

Create a `city_state` variable (e.g. “Baltimore, MD”) and then summarize within cities to obtain the total number of homicides and the number of unsolved homicides (those for which the disposition is “Closed without arrest” or “Open/No arrest”).

For the city of Baltimore, MD, use the prop.test function to estimate the proportion of homicides that are unsolved; save the output of prop.test as an R object, apply the broom::tidy to this object and pull the estimated proportion and confidence intervals from the resulting tidy dataframe.

```{r}
# Subset data for Baltimore
baltimore_data =
  homicides |> 
  filter(city_state == "Baltimore, MD")

# Apply prop.test
# Sum of unsolved cases and total number of cases
baltimore_data_unsolved = 
  baltimore_data |> 
  filter(disposition == "Closed without arrest" | disposition =="Open/No arrest")

prop_test_result = prop.test(nrow(baltimore_data_unsolved),nrow(baltimore_data))

# Use broom::tidy to tidy the prop.test result
tidy_result <- broom::tidy(prop_test_result)

# Extracting the estimated proportion and confidence intervals
estimated_proportion = 
  tidy_result |> 
  pull(estimate)
conf_low = 
  tidy_result |> 
  pull(conf.low)
conf_high = 
  tidy_result |> 
  pull(conf.high)

# Display the results
print(paste("Estimated proportion of unsolved cases: ", estimated_proportion))
print(paste("95% Confidence Interval: [", conf_low, ", ", conf_high, "]", sep = ""))
```



Now run prop.test for each of the cities in your dataset, and extract both the proportion of unsolved homicides and the confidence interval for each. Do this within a “tidy” pipeline, making use of purrr::map, purrr::map2, list columns and unnest as necessary to create a tidy dataframe with estimated proportions and CIs for each city.

```{r}
tidy_results = homicides |> 
  group_by(city) |> 
  nest() |> 
  mutate(
    prop_test = map(data, ~prop.test(
      sum(.x$disposition == "Closed without arrest" | .x$disposition == "Open/No arrest"), 
      nrow(.x)
    )),
    tidy_prop_test = map(prop_test, broom::tidy)
  ) |> 
  select(-data, -prop_test) |> 
  unnest(tidy_prop_test)
```


Create a plot that shows the estimates and CIs for each city – check out geom_errorbar for a way to add error bars based on the upper and lower limits. Organize cities according to the proportion of unsolved homicides.
```{r}
ggplot(tidy_results, aes(x = reorder(city, estimate), y = estimate)) +
  geom_point() +
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = 0.2) +
  coord_flip() +  # Flips the axes for better readability
  labs(x = "City", y = "Proportion of Unsolved Homicides", 
       title = "Proportion of Unsolved Homicides with Confidence Intervals by City") +
  theme_minimal()
```

Problem 2

This zip file contains data from a longitudinal study that included a control arm and an experimental arm. Data for each participant is included in a separate file, and file names include the subject ID and arm.

Create a tidy dataframe containing data from all participants, including the subject ID, arm, and observations over time
```{r}
file_names = list.files("data", full.names = TRUE)
data = map_df(file_names, read_csv, .id = "file_name")
```

```{r}
tidy_data <- data %>%
  mutate(
    subject_id = str_extract(file_names, "\\d+"),
    arm = str_extract(file_names, "con|exp")
  ) %>%
  select(-file_name) %>%
  pivot_longer(
    cols = -c(subject_id, arm),
    names_to = "time",
    values_to = "observation"
  )
```

Make a spaghetti plot showing observations on each subject over time, and comment on differences between groups.
```{r}
str(tidy_data)
```

```{r}
ggplot(tidy_data, aes(x = time, y = observation, color = subject_id, group = subject_id)) +
  geom_line() + 
  labs(title = "Observations Over Time by Subject and Group",
       x = "Time",
       y = "Observation Value",
       color = "Subject ID")+
  facet_grid(~arm)
```

Problem 3

When designing an experiment or analysis, a common question is whether it is likely that a true effect will be detected – put differently, whether a false null hypothesis will be rejected. The probability that a false null hypothesis is rejected is referred to as power, and it depends on several factors, including: the sample size; the effect size; and the error variance. In this problem, you will conduct a simulation to explore power in a one-sample t-test.

First set the following design elements:

Fix n=30
Fix σ=5
Set μ=0. Generate 5000 datasets from the model

x∼Normal[μ,σ]

```{r}
set.seed(123)  # For reproducibility
n <- 30        # Sample size
sigma <- 5     # Standard deviation
mu_values = 0:6

simulate_data <- function(mu, n, sigma) {
  x <- rnorm(n, mean = mu, sd = sigma)
  test_result <- broom::tidy(t.test(x, mu = 0))
  return(data.frame(mu = mu, estimate = mean(x), p_value = test_result$p.value))
}
```

```{r}
n_datasets <- 5000
all_results <- map_df(1:n_datasets, ~map_df(mu_values, ~simulate_data(.x, n, sigma), id = "mu"))
```

```{r}
all_results <- all_results %>%
  mutate(mu = as.numeric(mu),
         rejected_null = p_value < 0.05)
```

```{r}
all_results %>%
  group_by(mu) %>%
  summarize(power = mean(rejected_null)) |> 
  ggplot(aes(x = as.numeric(mu), y = power)) +
  geom_line() +
  labs(title = "Power vs. True Value of μ",
       x = "True Value of μ",
       y = "Power") +
  theme_minimal()
```


For each dataset, save μ̂ and the p-value arising from a test of H:μ=0 using α=0.05. Hint: to obtain the estimate and p-value, use broom::tidy to clean the output of t.test.

Repeat the above for μ={1,2,3,4,5,6}, and complete the following:

Make a plot showing the proportion of times the null was rejected (the power of the test) on the y axis and the true value of μ on the x axis. Describe the association between effect size and power.
Make a plot showing the average estimate of μ̂ on the y axis and the true value of μ on the x axis.

```{r}
all_results %>%
  group_by(mu) %>%
  summarize(mean_estimate = mean(estimate)) |> 
  ggplot(aes(x = as.numeric(mu), y = mean_estimate)) +
  geom_point() +
  labs(title = "Average Estimate vs. True Value of μ",
       x = "True Value of μ",
       y = "Average Estimate") +
  theme_minimal()
```

Make a second plot (or overlay on the first) the average estimate of μ̂ only in samples for which the null was rejected on the y axis and the true value of μ on the x axis. 

```{r}
all_results %>%
  group_by(mu) %>%
  filter(rejected_null == FALSE) |> 
  summarize(mean_estimate = mean(estimate)) |> 
  ggplot(aes(x = as.numeric(mu), y = mean_estimate)) +
  geom_point() +
  labs(title = "Average Estimate vs. True Value of μ",
       x = "True Value of μ",
       y = "Average Estimate") +
  theme_minimal()
```


Is the sample average of μ̂ across tests for which the null is rejected approximately equal to the true value of μ? Why or why not?
